### 2021-07-08 ch2flipFlop.tex
got the writing basically drafted. baout to change notation in the rwdifu chapter.

A relativity weak flow shearing a bed of sediment entrains individual particles into a state of motion controlled by turbulent forcing and intermittent collisions with other grains at rest on the bed, generating wide fluctuations in the sediment velocity \citep{Heyman2016,Fathel2015}.
Bed load particles move downstream until they are disentrained when they happen to encounter sufficiently sheltered pockets on the bed surface to interrupt their motions \citep{Charru2004,Gordon1972}.
Eventually, the bed around them rearranges and destroys this shelter, or turbulent fluctuations overcome the shelter \citep{Celik2014,Valyrakis2010}, particles are once again entrained, and the cycle repeats.
Bed load transport is thus a kind of itinerant motion, characterized by alternation between fluctuating movements and periods of rest.
This process is difficult to describe mathematically given the technical stochastic physics required \citep{Furbish2017,Ancey2020}.

@@

Many researchers have since refined Einstein's approach to provide more realistic descriptions of individual particle motions than Einstein's instantaneous step model.
One set of efforts has neglected the particle deposition process to calculate the downstream velocity distributions of moving particles after entrainment using mechanistic equations with terms representing turbulent fluid forces and particle-bed collisions \citep{Ancey2014,Fan2014}. 
Another set of works have described particles alternating between motion and rest, considering that motions occur with a constant velocity \citep{Lisle1998,Lajeunesse2017,Pierce2020b}. 
No models have yet been formulated which describe particles alternating between motion and rest with motions having realistic fluctuating velocities.

Parallel to these efforts to better describe individual particle trajectories, a complementary set of studies have developed stochastic formulations of the sediment flux which improve on the mean field description provided by Einstein \citep{Turowski2010,Furbish2012a,Ancey2020}. 
Sediment fluxes generally exhibit wide fluctuations due to variations in the number of moving particles and their velocities \citep{Bohm2005a,Ancey2006,Furbish2012a}, the migration of bedforms and sediment waves \citep{Guala2014,Recking2012}, and a host of other processes \citep{Dhont2018}.
Because these fluctuations occur over disparate timescales, measurements of mean sediment fluxes depend on the timescale over which they are collected, a phenomenon called scale-dependence \citep{Saletti2015,Dhont2018,Singh2009,Turowski2010,Ancey2020}.
To date, very few models have calculated the probability distribution of the bed load sediment flux due to individual particle motions \citep{Ancey2008,Ancey2014}, and among these, even fewer have described any observation-scale dependence of the flux \citep{Ancey2020,Turowski2010}.
Among those that do, none have yet formulated the sediment flux in terms of the downstream trajectories of individual grains.

This survey provides context for the two problems addressed in this chapter.
First, we do not yet have the capability to describe individual sediment trajectories through motion and rest including velocity fluctuations in the motion state; and second, we need more understanding of how to connect individual particle trajectories through motion and rest to the overall downstream sediment flux, including its probability distribution and the dependence of its statistical moments on the observation time.

Here, I develop a new statistical physics-based formalism which addresses both of these problems by describing individual particle trajectories with a stochastic Langevin equation.
This dynamical equation describes alternation between motion and rest at random intervals while particles in motion have velocity that fluctuates around a mean value.
Using the probability distribution of particle position generated by this model, I construct a formalism to derive analytically the probability distribution of the sediment flux.
This distribution exhibits observation-scale dependence as a result of velocity fluctuations among moving particles.
Below, I develop the new formalism in sec. \ref{sec:mod}, solve it in sec. \ref{sec:res}, and I discuss the implications of my results and future research ideas in secs. \ref{sec:disc} and \ref{sec:conc}.

\section{Description of motion-rest alternation with velocity fluctuations \label{sec:mod}}
The starting point for this analysis is an idealized one-dimensional domain, infinite in extent, populated with sediment particles on the surface of a sedimentary bed.
Particles are set in motion by the turbulent flow and move downstream until they deposit, and the cycle repeats.
The downstream coordinate is $x$, so that $\dot{x}$ describes a velocity in the downstream direction.

The flow is considered weak enough that interactions among moving grains are very rare, although interactions between moving particles and the bed may be common, characteristic of rarefied transport conditions \citep[e.g.][]{Kumaran2006,Furbish2017}. Any variations like particle size, shape, or spatial variations in the overlying fluid flow are neglected for mathematical simplicity.
These conditions allow for all particles to be described as independent from one another but governed by the same underlying dynamical equations.

\subsection{Dynamical equation for grain-scale sediment transport}
From these assumptions, the first target is to write an equation of motion for the individual sediment particle encompassing two features.
\begin{figure}
	\centerline{\includegraphics{./figures/ch2/fig0mod.pdf}}
	\caption{Panel (a) sketches a realization of the noise in equation \ref{eq:flippylangevin}, while panel (b) shows the trajectory derived from it (in blue) alongside other possible trajectories. Keys in panel (a) demonstrate the average movement time $1/k_D$ and rest time $1/k_E$, while the key in panel (b) shows the average movement velocity $V$. Velocity fluctuations lead to tilted stair-step trajectories with unsteady slopes in the $x-t$ plane. This can be compared with figure \ref{fig:lislefig} which does not include fluctuations.}
	\label{fig:fluxxy0}
\end{figure}
First, particles should alternate between motion and rest.
The transition rate from rest to motion is called entrainment and occurs with probability per unit time (or rate) $k_E$, while the transition from motion to rest is called deposition and occurs with rate $k_D$.
Second, particles in motion should move with mean velocity $V$ and their instantaneous velocities should jitter around this mean through time due to turbulent drag and particle-bed collisions.

The simplest equation of motion including these features is
\be \dot{x}(t) = [V + \sqrt{2D}\xi(t)]\eta(t).  \label{eq:flippylangevin} \ee
Here $\xi(t)$ is a Gaussian white noise having zero mean and unit variance representing velocity fluctuations among moving particles. $\eta(t)$ is a dichotomous noise which takes on values $\eta = 1$, representing motion (with mean duration $1/k_D$), and $\eta=0$, representing rest (with mean duration $1/k_E$).
Some trajectories produced by this equation are sketched in figure \ref{fig:fluxxy0}.

The mean particle velocity is $V$, and $D$ is a diffusivity [units $L^2/T$] describing velocity fluctuations among moving particles. The transition rate from $\eta=0$ to $\eta = 1$ is $k_E$, and the transition rate from $\eta=1$ to $\eta= 0$ is $k_D$. The notation $k=k_E+k_D$ is a shorthand used throughout the thesis.

The Langevin equation \ref{eq:flippylangevin} is a generalization of the model by \citet{Lisle1998} and \citet{Lajeunesse2017}, presented in section \ref{sec:lisle}, to include velocity fluctuations in the motion state.
It is similar to the ``randomly flashing diffusion" developed by Luczka and coworkers in physics \citep{Luczka1992,Luczka1993,Luczka1995}, although it includes a more complex asymmetric noise ($k_E\neq k_D$).
This equation describes the downstream movement of particles alternating through motion and rest with velocity fluctuations in the motion state. This is a more realistic description of bed load trajectories than earlier works in the Einstein paradigm reviewed in section \ref{sec:trajmodels}.

\subsection{Derivation of the master equation for the sediment position distribution}
\label{sec:floppymastereq}
The solution of equation \ref{eq:flippylangevin} for a given realization of the two noises $\eta(t)$ and $\xi(t)$ gives one possible trajectory of a particle. Averaging over the ensemble of all possible trajectories from different realizations of the noises will obtain the probability distribution $P(x,t)$ that a particle which started at position $x=0$ at time $t=0$ has travelled to position $x$ by time $t$. This distribution, by construction, will generalize earlier models of bedload alternation between motion and rest which did not include velocity fluctuations among moving particles \citep{Lisle1998,Lajeunesse2017}.

The desired probability distribution of position is formed as $ P(x',t) = \big\bra \delta(x'-x(t))\big\ket_{\eta,\xi} $, where $x(t)$ is the formal solution of eq. \ref{eq:flippylangevin} and the average is over both noises.
This symbolic equation is not directly useful as taking averages over both noises is a challenging mathematical problem \citep{Hanggi1978}. This calculation is completed in the appendix section \ref{sec:appAmaster}, providing the master equation
\be (\pt^2 + V \px \pt + k_E V \px + k \pt - D \px^2 \pt - k_E D \px^2) P(x,t) = 0 \label{eq:flippymaster}\ee
for the probability distribution of position.

The master equation \ref{eq:flippymaster} is a diffusion-like equation governing the probability distribution of position for individual particles as they transport downstream through a sequence of motions and rests, with the movement velocity being a fluctuating quantity.

One can see in particular that taking an the entrainment rate $k_E$ very large, meaning that all particles are generally moving, implies an advection-diffusion equation $(\pt + V\px -D \px^2)P=0$ for the position, characteristic of a particle moving downstream with Gaussian velocity fluctuations \citep{Ancey2014}. Otherwise, with $k_E$ of similar order as $k_D$, there is a finite probability that the particle is at rest, and the advection-diffusion process is often interrupted by deposition, giving rise to the additional terms in eq. \ref{eq:flippymaster} reminiscent of equation \ref{eq:lislemaster} from earlier simpler motion-rest models. When diffusion is turned off ($D=0$), equation \ref{eq:flippymaster} becomes identical to equation \ref{eq:lislemaster}, as expected.

@@

The probability distribution of the sediment flux can be calculated using the probability distribution of particle position $P(x,t)$ provided as the solution of equation \ref{eq:flippymaster}.
This method is modified from the approach recently developed by Banerjee and coworkers in physics \citep{Banerjee2020}. It generalizes the nonlocal formulation (\ref{sec:nonlocal}) to provide the probability distribution of the flux, and it generalizes the renewal approach (sec. \ref{sec:renewal}) to include the mechanics of individual particles in downstream transport.

@@

From this initial configuration, particles move downstream through time according to equation \ref{eq:flippylangevin}.
The flux is calculated as the average rate of particles crossing to the right of $x=0$ after a sampling time $T$.
\begin{figure}
	\centerline{\includegraphics{./figures/ch2/figure1.pdf}}
	\caption{The left panel indicates the configuration for the flux. The particle trajectories within are calculated from equation \ref{eq:langevin}, demonstrating alternation between rest and motion with fluctuating velocity. Particles begin their transport with positions $-L\leq x \leq 0$ at $t=0$, and as depicted in the right panel, the flux is calculated with the number of particles $N_>(T)$ which lie to the right of $x=0$ at the observation time $t=T$, divided by $T$: $q(T) = N_>(T)/T$. The probability distribution of $q(T)$ is determined from all possible realizations of the trajectories and initial positions as $N,L$ tend to infinity while the density of particles $\rho=N/L$ to the left of the control surface remains fixed.}
	\label{fig:flipflopfig1}
\end{figure}
This time-averaged flux is
\be q(T) = \frac{1}{T}\sum_{i=1}^N I_i(T). \label{eq:flippyflux} \ee
In this equation, $I_i(T)$ is an indicator function which is $1$ if the $i$th particle has passed the control surface ($x=0$) at the observation time $T$ and $0$ otherwise.
All particles which have not crossed the control surface (or which have crossed and then crossed back) contribute nothing to the flux.

The probability distribution of the flux is then the average of equation \ref{eq:flippyflux} across all possible initial configurations of particles and their trajectories
\be P(q|T) = \Big \bra \delta\Big( q - \frac{1}{T}\sum_{i=1}^N I_i(T) \Big) \Big\ket. \ee
To evaluate these averages, taking the Laplace transform over $q$ (i.e. forming the characteristic function of $P(q|T)$) obtains
\begin{align} \tilde{P}(s|T) &=  \Big\bra \exp\Big(\frac{s}{T}\sum_{i=1}^N I_i(T)\Big)\Big\ket \\
	&=  \prod_{i=1}^N \Big\bra \exp\Big(-\frac{s}{T}I_i(T)\Big)\Big\ket \\
	&= \prod_{i=1}^N \Big[1-\big(1-e^{-s/T}\big)\big\bra I_i(T) \big\ket \Big] \label{eq:flippycharacteristic}\end{align}
This progression relies on the independence of averages for each particle (so the average of a product is the product of averages) and the fact that  $e^{a I} = 1-(1-e^a)I$ because $I$ is either $0$ or $1$.

The average over initial conditions and possible trajectories for the $i$th particle involved in this characteristic function can be written
\be \bra I_i(t) \ket = \frac{1}{L}\int_L^0 dx' \int_0^\infty dx P(x-x',t) \label{eq:indicator} \ee
where $P(x,t)$ solves equation \ref{eq:flippymaster}.
Equation \ref{eq:indicator} describes the probability that the $ith$ particle is found to the right of control surface by time $T$ provided it started at a random location somewhere to the left.
These $\bra I_i(t) \ket$ are the components of the flux that depend on the particle dynamics.

Inserting equation \ref{eq:indicator} into equation \ref{eq:flippycharacteristic} and taking the limits $L\rightarrow \infty$ and $N \rightarrow \infty$ as the density of particles $\rho = N/L$ remains constant, obtaining the configuration of the nonlocal formulation in section \ref{sec:nonlocal}, provides
\begin{align} \tilde{P}(s|T) &= \lim_{N \rightarrow \infty} \Big(1 - \frac{1}{N}\big(1-e^{-s/T}\big)\Lambda(T) \Big)^N \\ &= \exp \Big[ -\big(1-e^{-s/T}\big)\Lambda(T) \Big]. \label{eq:flippychar} \end{align}
where $\Lambda(T) = \rho \int_0^\infty dx \int_0^\infty dx' P(x+x',T)$ is a rate constant, similar to the rate constant in the renewal process models of section \ref{sec:renewal}, except that it is now formulated in terms of individual particle trajectories via the probability distribution of particle position $P(x,t)$, which itself originates from the Langevin equation \ref{eq:flippylangevin} governing the particle dynamics.

The equation \ref{eq:flippychar} is the characteristic function of a Poisson distribution \citep{Cox1965}.
Expanding in $e^{-s/T}$ and inverting the Laplace transform provides a key equation of this chapter, the probability distribution of the flux, held contingent on the sampling time $T$:
\be P(q|T) = \sum_{l=0}^\infty \frac{\Lambda(T)^l}{l!}e^{-\Lambda(T)}\delta(q-\frac{l}{T}). \label{eq:flippydist}\ee
This equation implies that the mean flux is $\bra q(T) \ket = \int_0^\infty q P(q|T) dq = \Lambda(T)/T$, and similarly the variance is $\sigma_q^2(T) = \Lambda(T)/T^2$. For the case when $\Lambda(T)$ is proportional to the observation time ($\Lambda \propto T$), this equation is identical to the renewal theory approach reviewed in section \ref{sec:renewal}.
This demonstrates that the renewal approach can be formulated with more generality by considering the dynamics of individual particles as a starting point.

\section{Results \label{sec:res}}
\subsection{Position probability distribution of sediment particles}
\begin{figure}
	\centerline{\includegraphics{figures/ch2/figure2_slopeKey.pdf}}
	\caption{Panel (a) indicates the probability distribution of particle position (eq. \ref{eq:monster}) as it evolves through time. From the initial mixture of motion and rest states, particles advect downstream as they diffuse apart from one another due to differences in their velocities and exchange times between motion and rest. Panel (b) shows the resulting particle diffusion (eq. \ref{eq:diffo}). At timescales $t \ll 2D/V^2$, the diffusion is normal since the movement is approximately a standard Brownian diffusion process (as advection is irrelevant on these timescales). For larger timescales, $2D/V^2 \ll t \ll \tau$, particles undergo ballistic diffusion similar to \citet{Lisle1998} as a result of some particles being stationary as others advect. Finally at times longer than the timescale $\tau = 1/k$ associated with entrainment and deposition, diffusion is again normal, formed of particles well-mixed among motion and rest states. All results are scaled by the mean hop length $l=V/k_D$ and the autocorrelation time $\tau=1/k$ of the motion/rest alternation. In both plots, the lines are analytical results while the points are the results of Monte Carlo simulations based on evaluating cumulative transition probabilities on a small timestep \citep[e.g.][]{Barik2006}.}
	\label{fig:flippyfig1}
\end{figure}
The master equation \ref{eq:flippymaster} describes the evolution of the probability distribution of position through time. Loosely speaking the solution of this equation is expected to be some combination of the \citet{Einstein1937} theory for particle transport (section \ref{sec:einwalk}) with the Gaussian solution of the advection-diffusion equation \citep{Morse1953a}.

@@

The master equation \ref{eq:flippymaster} is solved by transform calculus in section \ref{sec:fluccymastersol} of the appendices, providing
\begin{multline} P(x,t) = \big[-\varphi D\px^2 + V\varphi \px + k + \delta(t) +  \pt \big]\\
	\times \int_0^t \mathcal{I}_0\Big( 2 \sqrt{k_Ek_D u(t-u)}\Big) e^{-k_E(t-u)-k_D u} \\ \times \sqrt{ \frac{1}{4\pi D u}} \exp\Big[- \frac{(x-Vu)^2}{4Du}\Big] du, \label{eq:monster}
 \end{multline}
where $\varphi=k_D/k$ is the probability the particle starts at rest. This equation generalizes the earlier result \citep[e.g.][]{Lisle1998} for alternation between motions and rests, eq. \ref{eq:lisledist}, to include velocity fluctuations within the motion state. This is another key result of this chapter.

The distribution \ref{eq:monster} is shown evolving through time in figure \ref{fig:flippyfig1} panel (a), where the advection and diffusion characteristics of equation \ref{eq:flippymaster} are both clearly visible. The integral in equation \ref{eq:monster} encodes the earlier expectation of Einstein model-like behavior mixed with a Gaussian propagator, in that it convolves the modified Bessel probability density that the particle has been in motion for a period $u$ out of a time $t$ with the advection-diffusion propagator that a particle has travelled a distance $x$ in time $u$. The prefactors of this integral term can be understood as adapting this distribution to the initial conditions.

@@

Using results from this generalized model as an example, I demonstrated how to calculate the sediment flux probability distribution, phrasing earlier renewal theory approaches (sec. \ref{sec:renewal}) more directly in terms of the underlying particle dynamics \citep[e.g.][]{Turowski2010,Ancey2020}.

This method to calculate the sediment flux distribution can also be interpreted as a generalization of the nonlocal formulation (sec. \ref{sec:nonlocal}) due to \citet{Parker2000,Furbish2012a} to describe sediment flux distributions, rather than just mean values.

@@

Models which generalize the approach established in this chapter to include collective motions should be capable of producing wide sediment transport fluctuations, but their development will be a challenging next step for research. The present chapter only lays the basic groundwork.

\subsection{Measurement protocol for the sediment flux}

>>>>

A relativity weak flow shearing a bed of sediment entrains individual particles into a state of motion controlled by turbulent forcing and intermittent collisions with other grains at rest on the bed, generating wide fluctuations in the sediment velocity \citep{Heyman2016,Fathel2015}.
Bed load particles move downstream until they are disentrained when they happen to encounter sufficiently sheltered pockets on the bed surface to interrupt their motions \citep{Charru2004,Gordon1972}.
Eventually, the bed around them rearranges and destroys this shelter, or turbulent fluctuations overcome the shelter \citep{Celik2014,Valyrakis2010}, and particles are once again entrained.
Bed load transport is thus a kind of itinerant motion, characterized by alternation between fluctuating movements and periods of rest.
This process is difficult to describe mathematically given the technicality of the required physics.

@@

Many researchers have since refined Einstein's approach to provide more realistic descriptions of individual particle motions than his instantaneous step model.
One set of efforts has neglected the particle deposition process to calculate the downstream velocity distributions of moving particles using mechanistic equations with noise terms representing turbulent fluid forces and particle-bed collisions \citep{Ancey2014,Fan2014}. 
Another set of works have described particles alternating between motion and rest considering that motions occur with a constant velocity \citep{Lisle1998,Lajeunesse2017}. 
No models have yet been formulated which describes particles alternating between motion and rest with motions having realistic fluctuating velocities, even though this is how bedload particles actually move.

Parallel to these efforts to better describe individual particle trajectories, a complementary set of studies have developed stochastic formulations of the sediment flux which improve on the mean field description provided by Einstein \citep{Turowski2010,Furbish2012a,Ancey2020}. 
Sediment fluxes generally exhibit wide fluctuations due to variations in the number of moving particles and their velocities \citep{Bohm2005a,Ancey2006,Furbish2012a}, the migration of bedforms and sediment waves \citep{Guala2014,Recking2012}, and a host of other processes \citep{Dhont2018}.
Because these fluctuations occur over disparate timescales, measurements of mean sediment fluxes depend on the timescale over which they are collected, a phenomenon called scale dependence \citep{Saletti2015,Dhont2018,Singh2009,Turowski2010,Ancey2020}.
To date, very few models have calculated the probability distribution of the bed load sediment flux due to individual particle motions \citep{Ancey2008,Ancey2014}, and among these, even fewer have described any observation-scale dependence of the flux \citep{Ancey2020a,Turowski2010}.
Among those that do, none have yet formulated the sediment flux in terms of the downstream trajectories of individual grains.

This survey provides context for the two problems addressed in this chapter.
First, we lack the capability to describe individual sediment trajectories through motion and rest including velocity fluctuations in the motion state; and second, we need more understanding of how to connect individual particle trajectories through motion and rest to the overall downstream sediment flux, including its probability distribution and the dependence of its statistical moments on the observation time.

Here, I develop a new statistical physics-based formalism which addresses both of these problems by describing individual particle trajectories with a stochastic Langevin equation.
This dynamical equation describes alternation between motion and rest at random intervals while particles in motion have velocity that fluctuates around a mean value.
Using the probability distribution of particle position generated by this model, I construct a formalism to derive analytically the probability distribution of the sediment flux.
This distribution exhibits observation-scale dependence as a result of velocity fluctuations among moving particles.
Below, I develop the new formalism in sec. \ref{sec:mod}, solve it in sec. \ref{sec:res}, and discuss the implications of my results and future research ideas in secs. \ref{sec:disc} and \ref{sec:conc}.

\section{Description of motion-rest alternation with velocity fluctuations \label{sec:mod}}
The starting point for this analysis is an idealized one-dimensional domain, infinite in extent, populated with sediment particles on the surface of a sedimentary bed.
Particles are set in motion by the turbulent flow and move downstream until they deposit, and the cycle repeats.
The downstream coordinate is $x$, so that $\dot{x}$ describes a velocity in the downstream direction.
The flow is considered weak enough that interactions among moving grains are very rare, although interactions between moving particles and the bed may be common, characteristic of rarefied transport conditions \citep[e.g.][]{Kumaran2006,Furbish2017}. 
Particles are considered to have similar enough shapes and sizes so as to have nearly identical mobility characteristics.
These conditions allow for all particles to be described as independent from one another but governed by the same underlying dynamical equations.

\subsection{Dynamical equation for grain-scale sediment transport}
From these assumptions, the first target is to write an equation of motion for the individual sediment particle encompassing two features.
\begin{figure}
	\centerline{\includegraphics{./figures/ch2/fig0mod.pdf}}
	\caption{Panel (a) sketches a realization of the noise in equation \ref{eq:flippylangevin}, while panel (b) shows the trajectory derived from it (in blue) alongside other possible trajectories. Keys in panel (a) demonstrate the average movement time $1/k_D$ and rest time $1/k_E$, while the key in panel (b) shows the average movement velocity $V$. Velocity fluctuations produce tilted stair-step trajectories with unsteady slopes in the $x-t$ plane. This can be compared with figure \ref{fig:lislefig} which does not include fluctuations.}
	\label{fig:fluxxy0}
\end{figure}
First, particles should alternate between motion and rest.
The entrainment transition rate from rest to motion occurs with probability per unit time (or rate) $k_E$, while the deposition transition from motion to rest occurs with rate $k_D$.
Second, particles in motion should move with mean velocity $V$ as their instantaneous velocities jitter around this mean value due to turbulent drag and particle-bed collisions.

The simplest equation of motion including these features is
\be \dot{x}(t) = [V + \sqrt{2D}\xi(t)]\eta(t).  \label{eq:flippylangevin} \ee
Here $\xi(t)$ is a Gaussian white noise having zero mean and unit variance representing velocity fluctuations among moving particles. $\eta(t)$ is a dichotomous noise which takes on values $\eta = 1$, representing motion (with mean duration $1/k_D$), and $\eta=0$, representing rest (with mean duration $1/k_E$).
Some trajectories produced by this equation are sketched in figure \ref{fig:fluxxy0}. In this equation, $V$ is the mean particle velocity describing the overall downstream drift of moving particles, while $D$ is a diffusivity [units $L^2/T^3$] describing velocity fluctuations among moving particles.
The transition rate from $\eta=0$ to $\eta = 1$ is $k_E$, and the transition rate from $\eta=1$ to $\eta= 0$ is $k_D$. The notation $k=k_E+k_D$ is a shorthand used throughout the thesis.

The Langevin equation \ref{eq:flippylangevin} is a generalization of the model by \citet{Lisle1998} and \citet{Lajeunesse2017}, presented in section \ref{sec:lisle}, to include velocity fluctuations in the motion stateprotocol.
It is similar to the ``randomly flashing diffusion" developed by Luczka and coworkers in physics \citep{Luczka1992,Luczka1993,Luczka1995}, although it is not clear if these models relate mathematically.
Equation \ref{eq:flippylangevin} describes the downstream movement of particles alternating through motion and rest with velocity fluctuations in the motion state. This dynamical equation provides a more realistic description of bed load trajectories than earlier works in the Einstein paradigm (sec. \ref{sec:trajmodels}).

\subsection{Derivation of the master equation for the sediment position distribution}
\label{sec:floppymastereq}
The solution of equation \ref{eq:flippylangevin} for a given realization of the two noises $\eta(t)$ and $\xi(t)$ gives one possible trajectory of a particle. The probability distribution $P(x,t)$ that a particle which started at position $x=0$ at time $t=0$ has travelled to position $x$ by time $t$ can be formulated as an average over the ensemble of all possible trajectories.

This probability distribution of position is formed as $ P(x',t) = \big\bra \delta(x'-x(t))\big\ket_{\eta,\xi} $, where $x(t)$ is the formal solution of eq. \ref{eq:flippylangevin} and the average is over both noises.
This symbolic equation is not directly useful as taking averages over both noises is a challenging mathematical problem \citep{Hanggi1978}. This calculation is completed in the appendix in section \ref{sec:appAmaster}, providing the master equation
\be (\pt^2 + V \px \pt + k_E V \px + k \pt - D \px^2 \pt - k_E D \px^2) P(x,t) = 0 \label{eq:flippymaster}\ee
for the probability distribution of position.
The master equation \ref{eq:flippymaster} is a diffusion-like equation governing the probability distribution of position for individual particles alternating between motion and rest, with the movement velocity being a fluctuating quantity.

One can see in particular that taking the entrainment rate $k_E$ very large, meaning that particles are very often moving, implies an advection-diffusion equation $(\pt + V\px -D \px^2)P=0$ for the position, characteristic of a particle moving downstream with Gaussian velocity fluctuations \citep{Ancey2014}. Otherwise, there is a possibility that the particle is at rest and the advection-diffusion process interrupted, giving rise to the additional terms in eq. \ref{eq:flippymaster}, a structure reminiscent of equation \ref{eq:lislemaster} from simpler motion-rest models.

@@

The probability distribution of the sediment flux can be calculated using the probability distribution of particle position $P(x,t)$ provided as the solution of equation \ref{eq:flippymaster}.
This method is modified from the approach recently developed by Banerjee and coworkers in physics \citep{Banerjee2020}.
This generalizes the nonlocal formulation (\ref{sec:nonlocal}) to provide the probability distribution of the flux, and it reframes the renewal approach (sec. \ref{sec:renewal}) to in terms of the mechanics of individual particles in transport.

@@

From this initial configuration, particles move downstream through time according to equation \ref{eq:flippylangevin}.
The flux is calculated as the average rate of particles crossing to the right of the control surface (at $x=0$) after the sampling time $T$.
\begin{figure}
	\centerline{\includegraphics{./figures/ch2/figure1.pdf}}
	\caption{The left panel indicates the configuration for the flux. The particle trajectories within are calculated from equation \ref{eq:langevin}, demonstrating alternation between rest and motion with fluctuating velocity. Particles begin their transport with positions $-L\leq x \leq 0$ at $t=0$, and as depicted in the right panel, the flux is calculated with the number of particles $N_>(T)$ which lie to the right of $x=0$ at the observation time $t=T$, divided by $T$: $q(T) = N_>(T)/T$. The probability distribution of $q(T)$ is determined from all possible realizations of the trajectories and initial positions as $N,L$ tend to infinity while the density of particles $\rho=N/L$ to the left of the control surface remains fixed.}
	\label{fig:flipflopfig1}
\end{figure}
This time-averaged flux is
\be q(T) = \frac{1}{T}\sum_{i=1}^N I_i(T). \label{eq:flippyflux} \ee
In this equation, the $I_i(T)$ are indicator functions which equal $1$ if the $i$th particle has passed the control surface ($x=0$) at the observation time $T$ and $0$ otherwise.
All particles which have not crossed the control surface (or which have crossed and then crossed back) contribute nothing to the flux.

The probability distribution of the flux is then the average of equation \ref{eq:flippyflux} across all possible initial configurations of particles and their trajectories
\be P(q|T) = \Big \bra \delta\Big( q - \frac{1}{T}\sum_{i=1}^N I_i(T) \Big) \Big\ket. \ee
These averages are again most easily conducted in Laplace space. 
Taking the Laplace transform over $q$ (i.e. forming the characteristic function of $P(q|T)$) obtains
\begin{align} \tilde{P}(s|T) &=  \Big\bra \exp\Big(\frac{s}{T}\sum_{i=1}^N I_i(T)\Big)\Big\ket \\
	&=  \prod_{i=1}^N \Big\bra \exp\Big(-\frac{s}{T}I_i(T)\Big)\Big\ket \\
	&= \prod_{i=1}^N \Big[1-\big(1-e^{-s/T}\big)\big\bra I_i(T) \big\ket \Big] \label{eq:flippycharacteristic}\end{align}
This progression relies on the independence of averages for each particle (so the average of a product is the product of averages) and the observation that  $e^{a I} = 1-(1-e^a)I$ because $I$ is either $0$ or $1$.

The average over initial conditions and possible trajectories for the $i$th particle involved in this characteristic function can be written
\be \bra I_i(t) \ket = \frac{1}{L}\int_L^0 dx' \int_0^\infty dx P(x-x',t) \label{eq:indicator} \ee
where $P(x,t)$ is the position probability distribution which solves equation \ref{eq:flippymaster}.
Equation \ref{eq:indicator} describes the probability that the $ith$ particle is found to the right of control surface by time $T$ provided it started at a random location somewhere to the left.
These $\bra I_i(t) \ket$ are the components of the flux that depend on the particle dynamics.

Inserting equation \ref{eq:indicator} into equation \ref{eq:flippycharacteristic} and taking the limits $L\rightarrow \infty$ and $N \rightarrow \infty$ as the density of particles $\rho = N/L$ remains constant, producing the configuration of the nonlocal formulation in section \ref{sec:nonlocal}, provides
\begin{align} \tilde{P}(s|T) &= \lim_{N \rightarrow \infty} \Big(1 - \frac{1}{N}\big(1-e^{-s/T}\big)\Lambda(T) \Big)^N \\ &= \exp \Big[ -\big(1-e^{-s/T}\big)\Lambda(T) \Big]. \label{eq:flippychar} \end{align}
where $\Lambda(T) = \rho \int_0^\infty dx \int_0^\infty dx' P(x+x',T)$ is a rate \textit{function}, similar to the rate constant in the renewal process models of section \ref{sec:renewal}, except that it is now formulated in terms of individual particle trajectories via the probability distribution of particle position $P(x,t)$, which itself originates from the Langevin equation \ref{eq:flippylangevin} governing the particle dynamics.

Equation \ref{eq:flippychar} is the characteristic function of a Poisson distribution \citep{Cox1965}.
Expanding in $e^{-s/T}$ and inverting the Laplace transform provides a key equation of this chapter, the probability distribution of the flux held contingent on the sampling time $T$:
\be P(q|T) = \sum_{l=0}^\infty \frac{\Lambda(T)^l}{l!}e^{-\Lambda(T)}\delta(q-\frac{l}{T}). \label{eq:flippydist}\ee
This equation implies that the mean flux is $\bra q(T) \ket = \int_0^\infty q P(q|T) dq = \Lambda(T)/T$. Similarly the variance is $\sigma_q^2(T) = \Lambda(T)/T^2$. For the case when $\Lambda(T)$ is proportional to the observation time ($\Lambda \propto T$), these formulas become identical to the renewal theory approach reviewed in section \ref{sec:renewal}.
This correspondence demonstrates that the renewal approach can be formulated equivalently by considering the dynamics of individual particles as a starting point.

\section{Results \label{sec:res}}
\subsection{Position probability distribution of sediment particles}
\begin{figure}
	\centerline{\includegraphics{figures/ch2/figure2_slopeKey.pdf}}
	\caption{Panel (a) indicates the probability distribution of particle position (eq. \ref{eq:monster}) as it evolves through time. From the initial mixture of motion and rest states, particles advect downstream as they diffuse apart from one another due to differences in their velocities and transition times between motion and rest. Panel (b) shows the resulting particle diffusion (eq. \ref{eq:diffo}). At timescales $t \ll 2D/V^2$, the diffusion is normal since the movement is approximately a standard Brownian diffusion process. For larger timescales, $2D/V^2 \ll t \ll \tau$, particles undergo ballistic diffusion similar to \citet{Lisle1998} as a result of some particles being stationary as others advect. Finally at times longer than the timescale $\tau = 1/k$ associated with entrainment and deposition, diffusion is again normal, formed of particles well-mixed among motion and rest states. All results are scaled by the mean hop length $\ell=V/k_D$ and the timescale $\tau=1/k$ of the motion/rest alternation. In both plots, the lines are analytical results while the points are the results of Monte Carlo simulations based on evaluating cumulative transition probabilities on a small timestep \citep[e.g.][]{Barik2006}.}
	\label{fig:flippyfig1}
\end{figure}
The master equation \ref{eq:flippymaster} describes the evolution of the probability distribution of position through time. The solution of this equation should be some combination of the \citet{Einstein1937} theory for particle transport (section \ref{sec:einwalk}) with the Gaussian solution of the advection-diffusion equation \citep{Morse1953a}.

@@

The master equation \ref{eq:flippymaster} is solved by transform calculus in section \ref{sec:fluccymastersol} of the appendix, providing
\begin{multline} P(x,t) = \big[-\varphi D\px^2 + V\varphi \px + k + \delta(t) +  \pt \big]\\
	\times \int_0^t \mathcal{I}_0\Big( 2 \sqrt{k_Ek_D u(t-u)}\Big) e^{-k_E(t-u)-k_D u} \\ \times \sqrt{ \frac{1}{4\pi D u}} \exp\Big[- \frac{(x-Vu)^2}{4Du}\Big] du, \label{eq:monster}
 \end{multline}
where $\varphi=k_D/k$ is the probability the particle starts at rest. This equation generalizes the earlier result  \ref{eq:lisledist} for alternation between motions and rests to include velocity fluctuations within the motion state.

The distribution \ref{eq:monster} is shown evolving through time in figure \ref{fig:flippyfig1} panel (a), where the advection and diffusion characteristics of equation \ref{eq:flippymaster} are both evident. The integral in equation \ref{eq:monster} encodes the earlier expectation of Einstein model-like behavior mixed with a Gaussian propagator, in that it convolves the Bessel probability that the particle has been in motion for a period $u$ out of a time $t$ with the Gaussian probability that a particle has travelled a distance $x$ in time $u$ within the motion state. The prefactors of this integral term can be understood as adapting this distribution to the initial conditions.

@@

Using results from this generalized model as an example, I demonstrated how to calculate the sediment flux probability distribution, phrasing earlier renewal theory approaches (sec. \ref{sec:renewal}) more directly in terms of the underlying particle dynamics \citep[e.g.][]{Turowski2010,Ancey2020}.
This method can also be viewed as a generalization of the nonlocal formulation (sec. \ref{sec:nonlocal}) developed by \citet{Parker2000,Furbish2012a} to describe sediment flux probability distributions, rather than just mean values.

@@

A generalization of the model developed in this chapter to include collective motions should be capable of producing realistically-wide sediment transport fluctuations. Such a development will be a challenging next step for research. The present chapter only lays the groundwork.

\subsection{Measuring the sediment flux}
